---
title: "Data Analysis in R"
author: "Simon Thornewill von Essen"
date: "12 12 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Now that we've covered the basics of R, we can start doing some basic data analysis.

Below, we have the scenario.

```
Suppose we are giving two students a multiple-choice exam w/ 40 questions where each question has 4 choices.

We don't know how much the students have studied for this exam, but we think that they will do better than just guessing at random.
```

Using these scenarios, we want to answer the following questions:

`1. What are the random paramters of interest?`

Here, we have two parameters. The likelihood of student 1 answering correctly and the likelihood of student 2 answering correctly.

`2. What is out likelihood?`

Hence, the likelihood is $\textrm{Binomial}(40, \theta)$.

`3. What prior should we use?`

The conjugate prior for a binomial distribution is a beta distribution.

`4. What is the prior prob P(thet1a>0.25)? P(theta1>0.5)? P(theta1>0.8)?`

```{r}
1-pbeta(0.25, 8, 4)
1-pbeta(0.50, 8, 4)
1-pbeta(0.80, 8, 4)

```

`5. Suppose that the first student gets 33 answers correct. What is the posterior distribution for theta2? P(thet1a>0.25)? P(theta1>0.5)? P(theta1>0.8)?`

```{r}
1-pbeta(0.25, 8+33, 4+40-33)
1-pbeta(0.50, 8+33, 4+40-33)
1-pbeta(0.80, 8+33, 4+40-33)

```

`6. Suppose the 2nd student gets 24 questions right. What is the posterior distribution for theta2 now?`

```{r}
correct <- 22
total <- 40

1-pbeta(0.25, 8+correct, 4+total-correct)
1-pbeta(0.50, 8+correct, 4+total-correct)
1-pbeta(0.80, 8+correct, 4+total-correct)

```

`7. What is the posterior probability that theta1>theta2, i.e., that the first student has a better chance of getting a question right htan the second student?`

```{r}
student_1.draws <- rbeta(shape1=8+33, shape2=4+40-33, n=1000)
student_2.draws <- rbeta(shape1=8+22, shape2=4+40-22, n=1000)

mean(student_1.draws > student_2.draws)

```
