---
title: "Introduction to R"
author: "Simon Thornewill von Essen"
date: "12 12 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

R is a statistical package.

When you type in math, it will try and solve this math.

```{r}
1+2*3
```

You can attach a name to the result of a calulation and save it to reference later.

```{r}
x <- 6  # Note you can use = instead of <- as well

x
```

You can create vectors using `c` to concatenate data.

```{r}
z <- c(2, 3, 4, 5, 6)

# You can also use the following
z <- c(2:6)

z
```

When we start having vectors (lists of numbers), we can start doing operations and aggregations on them. Note that R is not as much of an object oriented programming language, so instead of calling methods or attributes of an object, you usually use functions instead.

Knowing your functions is key to understanding and using R well.

```{r}
mean(z)
var(z)
```

You can also create sequences, like I userd above, using the `seq` function.

```{r}
seq(from=0.1, to=0.9, by=0.1)
```

If you're ever not sure what a function does, you can ask for help.

```{r}
help(seq)
```

R has some datasets which are already loaded.

```{r}
trees
```

We can attach the column names as variables so that we can call them individually.

```{r}
attach(trees)

Girth
```

We can now plot these vectors to get a feel for them.

```{r}
hist(Girth)
```

We can also look at scatter plots.

```{r}
plot(Girth, Height)
```

We can also look at matrix plots for the entire dataset.

```{r}
pairs(trees)
```

Finally, we can get an overview of the stats by calling `summary`.

```{r}
summary(trees)
```
