---
title: "Bayesian Linear Regression in R"
author: "Simon Thornewill von Essen"
date: "12/19/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

We can use linear regression in R in order to perform bayesian linear regression.

```{r}
oring <- read.table("http://www.randomservices.org/random/data/Challenger2.txt",
                    header = TRUE)

attach(oring)

head(oring)
```

The dataset that was just imported was the o-ring damage indices for the 
challenger 2 launch failure. The higher I is, the more damage was incurred on 
the ring.

```{r}
plot(T, I)
```

We can see that there is a rough linear relationship between the temperature, T,
and the damage index I. 

We can create a linear regression like so...

```{r}
oring.lm <- lm(I~T)

summary(oring.lm)
```

We can see here that for each increase in T we get a -.24 index of damage,
explaining 41% of the variance. 

## Given the data, what is the 95% posterior probability of the slope?

```{r}
-0.24337 - 0.06349*qt(0.975, 21)
-0.24337 + 0.06349*qt(0.975, 21)
```

This ends up becoming a posterior distribution because we multiply the error by
the prior, which is a t-distribution. 

Since this interval does not include zero, that should be good eviedence to
support the hypothesis that there is negative o-ring damage as the temperature
of the o-rings increases.

## For a given temp, whats the failure index

```{r}
18.36508 - 0.24337 * 31
```

However, sometimes we also would like an interval, we can do this like so...

```{r}
predict(oring.lm,
        data.frame(T=31),
        interval = "predict")
```

We can also calculate this manually. Note that we need to multiply by a scale
factor: 

$$\sqrt{1 + \frac{1}{n} + \frac{(x^*-\bar{x})^2}{(n-1)s_x^2}}$$

```{r}
10.82052 - 2.102*qt(0.975, 21)*sqrt(1 + 1/23 + ((31-mean(T))^2/22/var(T)))

10.82052 + 2.102*qt(0.975, 21)*sqrt(1 + 1/23 + ((31-mean(T))^2/22/var(T)))
```

We can even use this to calculate probabilistic statements of failure given a
certain point.

```{r}
1 - pt((0 - 10.82061)/(2.102*sqrt(1 + 1/length(T) +
                                    ((31-mean(T))^2/(length(T)-1)/var(T)))),
       21)
```

However, doing this analytically is kind of tedious, normally we have packages
that are able to calculate these things for us such as JAGS in R or PyMC/Pyro in
python.
