---
title: "Plotting Likelihood in R"
author: "Simon Thornewill von Essen"
date: "12 12 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this markdown file we will take a closer look at likelihood. Namely, a bernouli likelihood.

```{r}
likelihood <- function(n, y, theta){
  return(theta^y*(1-theta)^(n-y))
}
```

We can text our likelihood function out:

```{r}
theta <- seq(from=0.01, to=0.99, by=0.01)

plot(theta, likelihood(400, 72, theta))
```

If we look at the plot above, we can see that the probability of seeing 72 positives out of 400 flips is maximised at 18%.

We can add lines using the function `abline` to show that this is true.

```{r}
plot(theta, likelihood(400, 72, theta))
abline(v=.18)
```

This can also be made more numerically stable by using the log likelihood. A function for this is created below.

```{r}
loglike <- function(n, y, theta){
  return(y*log(theta)+(n-y)*log(1-theta))
}
```

Plotting the result, we get the following.

```{r}
plot(theta, loglike(400, 72, theta), type='l')
abline(v=.18)
```

