---
title: "Two Factor ANOVA"
author: "Simon Thornewill von Essen"
date: "29 1 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

set.seed(83)

library("rjags")
library("coda")
library("ggplot2")
```

```{r}
data("warpbreaks")
head(warpbreaks, 2)
```

For this dataset, we just the number of warp break per loom (fixed length of yarn). Columns correspond to the number of breaks, the type of wool (A or B) and the level of tension (Low, Medium, High).

We can do a pivot table on this data to see the number of observations for each kind of break.

```{r}
table(warpbreaks$wool, warpbreaks$tension)
```

We have 9 observations for each kind of break, this means that the sample has an even distribution of breaks, which is good.

```{r}

warpbreaks$intrctn <- interaction(warpbreaks$wool, warpbreaks$tension)

ggplot(data=warpbreaks, aes(x=intrctn, y=breaks, fill=intrctn)) + 
  geom_boxplot()
```

From this visualization, it looks like both factors may play a role in the number of breaks. It appears that there is a general decrease in breaks as we move from low to medium to high tension. Let’s start with a one-way model using tension only.

## One way Model

```{r}
mod1_string = " model {
    for( i in 1:length(y)) {
        y[i] ~ dnorm(mu[tensGrp[i]], prec)
    }
    
    for (j in 1:3) {
        mu[j] ~ dnorm(0.0, 1.0/1.0e6)
    }
    
    prec ~ dgamma(5/2.0, 5*2.0/2.0)
    sig = sqrt(1.0 / prec)
} "

data1_jags = list(y=log(warpbreaks$breaks), tensGrp=as.numeric(warpbreaks$tension))

params1 = c("mu", "sig")

mod1 = jags.model(textConnection(mod1_string), data=data1_jags, n.chains=3)
update(mod1, 1e3)

mod1_sim = coda.samples(model=mod1,
                        variable.names=params1,
                        n.iter=5e3)

## convergence diagnostics
plot(mod1_sim)
```

```{r}
summary(mod1_sim)
```

Here, we can see that the 95% posterior HPD for the mean of group 2 overlaps with the low and high groups, but the intervals for the other two only slighly overlap. This is a strong indication that the means for low and high tension are different.

```{r}
dic1 <- dic.samples(mod1, n.iter = 1e3)

dic1
```

## Two Way Model

With two factors, we have six treatment groups. We will first fir the additive model, which treats the two factors separately with no interaction. 

```{r}
# Create design matrix
X <- model.matrix(~ wool + tension, data=warpbreaks)
head(X)
```
```{r}
tail(X)
```

R has chosen the mean for wool A and low tension to be the intercept. 

Now, we can build our model.

```{r}
mod2_string <- "model{
  for(i in 1:length(y)){
    y[i] ~ dnorm(mu[i], prec)
    mu[i] = int + alpha*isWoolB[i] + beta[1]*isTensionM[i] + beta[2]*isTensionH[i]
  }
  
  int ~ dnorm(0.0, 1.0/1.0e6)
  alpha ~ dnorm(0.0, 1.0/1.0e6)
  
  for(j in 1:2){
    beta[j] ~ dnorm(0.0, 1.0/1.0e6)
  }
  
  prec ~ dgamma(3/2.0, 3*1.0/2.0)
  sig = sqrt(1.0 / prec)
}"

data2_jags <- list(y=log(warpbreaks$breaks), isWoolB=X[,"woolB"], isTensionM=X[,"tensionM"], isTensionH=X[,"tensionH"])

params2 <- c("int", "alpha", "beta", "sig")

mod2 <- jags.model(textConnection(mod2_string), data=data2_jags, n.chains=3)
update(mod2, 1e3)

mod2_sim <- coda.samples(model=mod2,
                        variable.names=params2,
                        n.iter=5e3)

plot(mod2_sim)
```

```{r}
summary(mod2_sim)
```

```{r}
dic2 <- dic.samples(mod2, n.iter=1e3)

dic1

dic2
```

We can see that the DIC has gone down a little bit by adding the wool factor to the model. But if we look at the aplot showing the breaks for string B, we can see that there are more breaks than string 1 when under moderate tension. Hence, it might be appropriate to add an interaction term.

```{r}
lmod2 <- lm(log(breaks) ~ .^2, data=warpbreaks)

summary(lmod2)
```

To model this in RJAGS

```{r}
mod3_string <- "model{
  for(i in 1:length(y)){
    y[i] ~ dnorm(mu[woolGrp[i], tensGrp[i]], prec)
  }
  
  for(j in 1:max(woolGrp)){
    for(k in 1:max(tensGrp)){
      mu[j, k] ~ dnorm(0.0, 1.0/1.0e6)
    }
  }
  prec ~ dgamma(3/2.0, 3*1.0/2.0)
  sig = sqrt(1.0 / prec)
}"

data3_jags <- list(y=log(warpbreaks$breaks), woolGrp=as.numeric(warpbreaks$wool), tensGrp=as.numeric(warpbreaks$tension))

params3 <- c("mu", "sig")

mod3 <- jags.model(textConnection(mod3_string), data=data3_jags, n.chains=3)
update(mod3, 1e3)

mod3_sim <- coda.samples(model=mod3,
                        variable.names=params3,
                        n.iter=5e3)
mod3_csim <- as.mcmc(do.call(rbind, mod3_sim))
```

```{r}
par(mfrow=c(3,2)) # arrange frame for plots
densplot(mod3_csim[,1:6], xlim=c(2.0, 4.5))
```

Note, that while it might be tempting to look at comparisons between the combination of treatments, but it might lead to false positives. It's best not to search results for interesting hypotheses, because if there are many hypotheses then some will appear to show associations simply due to random chance.

Results are most reliable when we determine a relatively small number of hypotheses we are interested in beforehand, collect the data and then statistically evaluate the evidence for them.

One question we might be interested in is finding the treatment combo that produces the fewest breaks. To calculate this, we can go through our posterior samples and for each simple, fine out which group has the smallest mean. 

```{r}
prop.table( table( apply(mod3_csim[,1:6], 1, which.min) ) )
```

The evidence supports wool B with high tension as the treatment that produces the fewest breaks.
